Constraints on Assembly Bias from Galaxy Clustering

### 2016-06-17 ms.tex
Added Concluding Paragraph.

\maketitle
%----------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%  A B S T R A C T %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  We constrain the newly-introduced decorated Halo Occupation Distribution (HOD) 
  model using SDSS DR7 measurements of projected galaxy clustering, $\wprp (r_{\rm p})$ 
  of galaxies in r-band luminosity-threshold samples. 
  The decorated HOD is a model for the galaxy--halo
  connection that augments the traditional HOD by allowing for the possibility of 
  {\em assembly bias}: galaxy luminosity may be correlated with dark matter halo
  properties besides mass, $\mvir$. We demonstrate that it
  is not possible to rule out galaxy assembly bias using DR7 measurements of galaxy 
  clustering alone. Moreover, galaxy samples $\magr<-20, -20.5$
  favor strong levels of central galaxy assembly bias. These samples favor models 
  in which high-concentration halos are more likely to host a central galaxy
  relative to low-concentration halos of the same $\mvir.$ We rule out
  zero assembly bias with high significance for these
  samples. Satellite galaxy assembly bias is significant for the
  faintest sample we study, $\magr<-19.$ We find no evidence for
  assembly bias in the $\magr<-21$ sample. In addition to presenting the first 
  constraints on HOD models that accommodate assembly bias, our analysis 
  includes numerous improvements over previous analyses of this data set and 
  supersedes previously published results, even in the case of a standard HOD analysis.
\end{abstract}

@@

@@

We have re-analyzed the SDSS DR7 measurements of projected galaxy clustering,
$\wprp,$ and number density, $n_{\rm g},$ originally published in \citet{zehavi_etal11}.
The novelty of our work is that we have performed these analyses using the
decorated Halo Occupation Distribution (HOD) formalism, which allows for galaxy
assembly bias and was introduced by \citet{hearin_etal16} precisely for this purpose.
In this work, we provide the first quantitative constraints on
assembly bias derived from the Decorated HOD. 
We enumerate our most important conclusions below.
%
\ben
%
\item It is not possible to rule out galaxy assembly bias using SDSS DR7 measurements of $\wprp$ and $n_{\rm g}.$
%
\item Decorated HOD fits to $\wprp$ and $n_{\rm g}$ favor significant levels of assembly bias, particularly in the lower luminosity thresholds we study. Both the $\magr<-20, -20.5$ samples prefer relatively strong central galaxy assembly bias. The $\magr<-19$ sample favors satellite assembly bias. Assembly bias in these samples improves the capability of the model to describe galaxy clustering near
the one-halo to two-halo transition at $r_{\rm p} \sim 2 \, h^{-1}\mathrm{Mpc}$.
%
\item The evidence for galaxy assembly bias weakens for brighter galaxy samples: Decorated HOD fits to the $\magr<-21$ sample are consistent with both $A_{\rm cens}=0$ and $A_{\rm sats}=0.$ This is consistent with the well-established result that
{\em halo assembly bias} weakens with increasing halo mass over the dynamic range relevant to these galaxy samples
\citep[see, e.g., Figure 8 of][and references therein]{hearin_etal16}.
%
\item Our posteriors and best-fit parameters summarized in
Table \ref{table:parameters} supersede the values published in \citet{zehavi_etal11},
as direct-mock-population together with the Decorated HOD allows us to account for
highly significant systematics that have heretofore been neglected from all HOD fits to SDSS data.
We note that our findings update the original \citet{zehavi_etal11} results
{\em even for our fits to standard HODs without assembly bias}, because
the original results were derived from MCMC chains that did not sufficiently sample the HOD model posteriors.
%
\een
%
We conclude by noting that since $\wprp$ and $n_{\rm g}$ are already very well-measured in DR7,
it is likely that further improvements on assembly bias constraints at low-redshift will require additional
observational measurements, e.g., galaxy--galaxy lensing, group statistics, etc.
I need to think of a more clever ending.

>>>>

\maketitle
%----------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%  A B S T R A C T %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  We constrain the newly-introduced decorated Halo Occupation Distribution (HOD) 
  model using SDSS DR7 measurements of projected galaxy clustering, $\wprp (r_{\rm p})$ 
  of galaxies in r-band luminosity-threshold samples. 
  The decorated HOD is a model for the galaxy--halo
  connection that augments the traditional HOD by allowing for the possibility of 
  {\em assembly bias}: galaxy luminosity may be correlated with dark matter halo
  properties besides mass, $\mvir$. We demonstrate that it
  is not possible to rule out galaxy assembly bias using DR7 measurements of galaxy 
  clustering alone. Moreover, galaxy samples $\magr<-20, -20.5$
  favor strong levels of central galaxy assembly bias. These samples favor models 
  in which high-concentration halos are more likely to host a central galaxy
  relative to low-concentration halos of the same $\mvir.$ We rule out
  zero assembly bias with high significance for these
  samples. Satellite galaxy assembly bias is significant for the
  faintest sample we study, $\magr<-19.$ We find no evidence for
  assembly bias in the $\magr<-21$ sample. Assembly bias should be accounted for 
  in galaxy clustering analyses or attempts to exploit galaxy clustering to constrain cosmology. 
  In addition to presenting the first constraints on HOD models that accommodate assembly bias, our analysis 
  includes numerous improvements over previous analyses of this data set and 
  supersedes previously published results, even in the case of a standard HOD analysis.
\end{abstract}

@@

%------------------------------------------------
\section{Interpretation}
\label{section:discussion}
%------------------------------------------------

@@

We have re-analyzed the SDSS DR7 measurements of projected galaxy clustering,
$\wprp,$ and number density, $n_{\rm g},$ originally published in \citet{zehavi_etal11}.
The novelty of our work is that we have performed these analyses using the
decorated Halo Occupation Distribution (HOD) formalism, which allows for galaxy
assembly bias and was introduced by \citet{hearin_etal16} precisely for this purpose.
In this work, we provide the first quantitative constraints on
assembly bias derived from the Decorated HOD. Projected galaxy clustering, 
$\wprp(r_{\rm p})$ is already very well-measured within SDSS DR7, therefore it is likely 
that further improvements on assembly bias constraints at low-redshift will require the use of additional 
observables such as galaxy--galaxy lensing, group statistics, void statistics, or numerous other possibilities. 
We are currently exploring many of those possibilities and will report on those results in a forthcoming publication.

Our most important conclusions are as follows.
%
\ben
%
\item It is not possible to rule out galaxy assembly bias using SDSS DR7 measurements of $\wprp$ and $n_{\rm g}.$
%
\item Decorated HOD fits to $\wprp$ and $n_{\rm g}$ favor significant levels of assembly bias, particularly in the lower luminosity thresholds we study. Both the $\magr<-20, -20.5$ samples prefer relatively strong central galaxy assembly bias. The $\magr<-19$ sample favors satellite assembly bias. Assembly bias in these samples improves the capability of the model to describe galaxy clustering near
the one-halo to two-halo transition at $r_{\rm p} \sim 2 \, h^{-1}\mathrm{Mpc}$.
%
\item The evidence for galaxy assembly bias weakens for brighter galaxy samples: Decorated HOD fits to the $\magr<-21$ sample are consistent with both $A_{\rm cens}=0$ and $A_{\rm sats}=0.$ This is consistent with the well-established result that
{\em halo assembly bias} weakens with increasing halo mass over the dynamic range relevant to these galaxy samples
\citep[see, e.g., Figure 8 of][and references therein]{hearin_etal16}.
%
\item Our posteriors and best-fit parameters summarized in
Table \ref{table:parameters} supersede the values published in \citet{zehavi_etal11},
as direct-mock-population together with the Decorated HOD allows us to account for
highly significant systematics that have heretofore been neglected from all HOD fits to SDSS data.
We note that our findings update the original \citet{zehavi_etal11} results
{\em even for our fits to standard HODs without assembly bias}, because
the original results were derived from MCMC chains that did not sufficiently sample the HOD model posteriors.
%
\een
%

Until such time as we have stringent constraints on the level of assembly bias in the 
universe, accounting for the possibility of assembly bias is not optional in interpreting 
galaxy clustering data. Precise statements about galaxy evolution derived from 
galaxy clustering or attempts to exploit clustering to perform cosmological analyses 
cannot be made robustly without accounting for assembly bias. Together with 
the work of \citet{hearin_etal16} and the {\tt Halotools} software, we have provided 
an explicit example of one way forward in such analyses. It is our hope that similar analyses 
will be performed on forthcoming galaxy survey data and that these analyses will lead to a
richer understanding of galaxy evolution and the relationship between galaxies and the halos 
and environments in which they are found.
